\section{bidirectional\_\-hds\_\-concept.hpp File Reference}
\label{bidirectional__hds__concept_8hpp}\index{bidirectional_hds_concept.hpp@{bidirectional\_\-hds\_\-concept.hpp}}
Bidirectional HDS concept definition. 



\subsection{Detailed Description}
Bidirectional HDS concept definition. 

\begin{Desc}
\item[Definition]This concept refines the Forward\-HDSConcept and Backward\-HDSConcept, and specifies additional bidirectional traversal category. Bidirectional category requires both the forward\_\-category and the backward\_\-category defined to iterate over the HDS data structure, which means both clockwise \& counter clockwise movement around the facet and vertex. From these two categories all others can be accessed with the help of the opposite of a halfedge.\end{Desc}
\begin{Desc}
\item[Refinement of]\begin{itemize}
\item Forward\-HDSConcept\item Backward\-HDSConcept\end{itemize}
\end{Desc}
\begin{Desc}
\item[Notation]\begin{itemize}
\item HDS A type that is a model of Bidirectional\-HDSConcept\item hds A non-modifiable instance of HDS\item h,g Halfedge descriptors, of type X::halfedge\_\-descriptor\end{itemize}
\end{Desc}
\begin{Desc}
\item[Associated types]\begin{itemize}
\item hds\_\-traits$<$HDS$>$::halfedge\_\-descriptor: must be Default\-Constructible and Copy\-Constructible, Equality\-Comparable, Assignable.\item hds\_\-traits$<$HDS$>$::traversal\_\-category: must be convertible to hds\_\-traits$<$HDS$>$::bidirectional\_\-traversal\_\-tag.\end{itemize}
\end{Desc}
\begin{Desc}
\item[Definitions]\begin{itemize}
\item {\em halfedge\/} {\em descriptor\/} is a type that contains information to access the halfedge.\item {\em traversal\/} {\em category\/} is a type that contains information about the direction of the traversal, which is forward, backward or bidirectional. For bidirectional hds concept, it has to be convertible to hds\_\-traits$<$HDS$>$::bidirectional\_\-traversal\_\-tag.\end{itemize}
\end{Desc}
\begin{Desc}
\item[Valid Expressions]\begin{itemize}
\item see Forward\-HDSConcept and Backward\-HDSConcept.\end{itemize}
\end{Desc}
\begin{Desc}
\item[Expression Semantics]\begin{itemize}
\item see Forward\-HDSConcept and Backward\-HDSConcept.\end{itemize}
\end{Desc}
\begin{Desc}
\item[Complexity guarantees]\begin{itemize}
\item see Forward\-HDSConcept and Backward\-HDSConcept.\end{itemize}
\end{Desc}
\begin{Desc}
\item[Invariants ]The invariants for Forward\-HDSConcept, and Backward\-HDSConcept should hold for Bidirectional\-HDSConcept also.\end{Desc}
\begin{Desc}
\item[Concept-checking class]

\begin{Code}\begin{verbatim} template <class HDS> 
 struct BidirectionalHDSConcept {
   typedef typename hds_traits<HDS>::halfedge_descriptor halfedge_descriptor; 
   typedef typename hds_traits<HDS>::traversal_category traversal_category;
    void constraints() {
       using namespace boost;
       function_requires<ForwardHDSConcept<HDS> >();
       function_requires<BackwardHDSConcept<HDS> >();
       function_requires<ConvertibleConcept<traversal_category,
                                          bidirectional_traversal_tag> >();  
    }
    BidirectionalHDS hds;
 };
\end{verbatim}\end{Code}

 \end{Desc}
